---
title: "In Class Review"
author: "Dr. Shoemaker"
date: "DATA 2401"
output:
  html_document: default
  pdf_document: default
---

**Note: Your exam will be in a similar format to this (RMarkdown File with Questions and Code chunks for your answers), so make SURE you have the process for knitting to html or pdf working.** 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# uncomment these, run them ONCE, then recomment them
# install.packages(tidyverse)
# install.packages(ggplot2)


library(tidyverse)
library(ggplot2)

```

## Indexing 

1. The below code chunk saves some built-in data about diamonds to vectors in your environment. (Run `? diamonds` in your console for more information, you may need to install and load the `tidyverse` and `ggplot2` libraries using the code in the chunk above). Using indexing and the tools you've learned, find the following:

   - What's the mean price of diamonds? The mean carat value? 
   - How many diamonds are more than 4 carats?
   - What's the carat weight of the largest diamond?
   - What's the price of most expensive diamond? 
   - Assuming that these are paired information (`carats[1]` and `price[1]` are information about the *same diamond*), how many carats is the most expensive diamond? How expensive is the biggest diamond? 
      + I've added some extra questions below for the types of thing you can answer. 

```{r}
# I'm taking these out of the included data dataframe so we can work with them as their own vectors. 
carats <- diamonds$carat
price <- diamonds$price

```

```{r}
# What's the mean price of diamonds? The mean carat value? 


# How many diamonds are more than 4 carats?


# What's the largest diamond?


# What's the most expensive diamond? 


# Assuming that these are paired information 
# (`carats[1]` and `price[1]` are information about the same diamond), 
# how many carats is the most expensive diamond? 


# find average price of diamonds under 1 carat

# what is size of the largest diamond I can get for less than 1000

```

##  Functions + Indexing

1. Write a function that takes in a numeric vector, changes any negative values to 0, and returns the changed vector. 

```{r}


```






##  Lists

Back to the diamond data. 

1. Make a boolean vector called `too_much` that is T/F to indicate if each element of `price` is larger than the median value of `price`.

```{r}


```


1. Put this vector, along with `carats` and `price` in a list called `diamond_info`. What are the different ways you can access the information inside this list? 

```{r}

```

```{r, eval = F}
# I set eval = F here in the chunk so these don't run and take 200+ pages... 

```

```{r}
# let's look at what happens here
# back to eval = T, which is the default


```


1. Use `lapply` to find the mean of each variable. 
Sidenote challenge question: why is the mean value of `too_much` what it is? 

```{r}


```

1. How can I remove the element `too_much` from the list?

```{r}

```



```{r}
# price of the diamonds that are larger than 4 carats


# price of the diamonds that are larger than 4 carats or smaller than .3 carats

```

