---
title: "In Class Review"
author: "Dr. Shoemaker"
date: "DATA 2401"
output:
  html_document: default
  pdf_document: default
---

**Note: Your exam will be in a similar format to this (RMarkdown File with Questions and Code chunks for your answers), so make SURE you have the process for knitting to html or pdf working.** 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# uncomment these, run them ONCE, then recomment them
# install.packages(tidyverse)
# install.packages(ggplot2)


library(tidyverse)
library(ggplot2)

```

## Indexing 

1. The below code chunk saves some built-in data about diamonds to vectors in your environment. (Run `? diamonds` in your console for more information, you may need to install and load the `tidyverse` and `ggplot2` libraries using the code in the chunk above). Using indexing and the tools you've learned, find the following:

   - What's the mean price of diamonds? The mean carat value? 
   - How many diamonds are more than 4 carats?
   - What's the carat weight of the largest diamond?
   - What's the price of most expensive diamond? 
   - Assuming that these are paired information (`carats[1]` and `price[1]` are information about the *same diamond*), how many carats is the most expensive diamond? How expensive is the biggest diamond? 
      + I've added some extra questions below for the types of thing you can answer. 

```{r}
# I'm taking these out of the included data dataframe so we can work with them as their own vectors. 
carats <- diamonds$carat
price <- diamonds$price

```

```{r}
# What's the mean price of diamonds? The mean carat value? 
mean(price)

# How many diamonds are more than 4 carats?
sum(carats > 4)
length(carats[carats > 4])
table(carats > 4)

# What's the largest diamond?
max(carats)


# What's the most expensive diamond? 
max(price)
min(price) # least expensive

# Assuming that these are paired information 
# (`carats[1]` and `price[1]` are information about the same diamond), 
# how many carats is the most expensive diamond? 

## find the index value for the most expensive, use that in the carats vector
price[price == max(price)] ## same value as max(price)
which.max(price) # is 27750
price[27750]
price[which.max(price)]
carats[which.max(price)]

price[which.max(carats)]
price[which.min(carats)]

# find average price of diamonds under 1 carat

# what is size of the largest diamond I can get for less than 1000

```

##  Functions + Indexing

1. Write a function that takes in a numeric vector, changes any negative values to 0, and returns the changed vector. 

```{r}
x <- 2:-4
x[x < 0] <- 0

myfun <- function(orange) {
   min(orange) + max(orange)
}
   
myfun(4:-2)
```




##  Lists

Back to the diamond data. 

1. Make a boolean vector called `too_much` that is T/F to indicate if each element of `price` is larger than the median value of `price`.

```{r}
too_much <- price > median(price)

```


1. Put this vector, along with `carats` and `price` in a list called `diamond_info`. What are the different ways you can access the information inside this list? 

```{r}
diamond_info <- list(too_much, price, carats)
# diamond_info[[2]] 

diamond_info <- list(my_too_much = too_much, 
                     my_price = price, 
                     my_carats = carats)

diamond_info$my_price[which.max(diamond_info$my_carats)]
```


1. Use `lapply` to find the mean of each variable. 
Sidenote challenge question: why is the mean value of `too_much` what it is? 

```{r}
lapply(diamond_info, mean)

```

1. How can I remove the element `too_much` from the list?

```{r}
diamond_info$my_too_much <- NULL
diamond_info$color <- c("clear", "yellow")
str(diamond_info)
```



```{r}
# price of the diamonds that are smaller than 1 carats
diamond_info[['my_price']][diamond_info$my_carats < 1]
diamond_info$my_price[diamond_info$my_carats < 1]



# price of the diamonds that are larger than 4 carats or smaller than .3 carats

## | is OR 
## & is AND

diamond_info$my_price[diamond_info$my_carats > 4 | diamond_info$my_carats < .3]

## function that gives mean price of diamonds under size X

smaller <- function(X){
   mean(diamond_info$my_price[diamond_info$my_carats < X])
}

smaller(4)
```




## In class work 

```{r}
# rbind cbind
a <- matrix(1:6, nrow = 2)
b <- matrix(11:16, nrow = 2)


# if else 

x <- 1:6

if (x > 2) { 
   x <- x - 1 
}  else {
   x <- x + 34
}
   

# Indexing
# we can index by number (hard coding), by vector, by conditional statement, or by name list[['name of element']]

# Functions

### writing it
my_fun <- function(var1 = 9, var2 = 2){
   (x <- var1 + var2)
}

### using it
my_fun(var2 = 8)
my_fun()


# Vectors vs lists


# Is.list, Is.na, is.vector, is.numeric, is.XXXXX
### helper functions
y <- 1:6
y[100] <- "b"
y
sum(is.na(y))
x <- 1:6
sum(x > 4)

x <- c(T, F, T)
sum(x)
   
# The boolean stuff (all, xor, etc) and the data type you can use them on
# also helper function, but for logicals
z <- 1:7
any(is.na(z))
xor(T, T)

# Single vs. Double brackets
   # list 
# Dollar Sign notation
   # list
# Lapply

# Recycling
x <- 1:6
y <- 1:10
x + y 
```