---
title: "Review_day_Exam1"
author: "Nhan Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
### R review before test 1- do the test 1 sample online 
#### Stupid Dee all the way up


Part 1 (12 points)
1. Create the vector of all the integers from -10 to 10 in three ways: once using c(), once using :, and
once using seq()
2. Create the vector [3, 6, 9, 12, 15] in two ways, once using c() and once using seq()
3. Create a vector that contains the log of the integers from 15 to 25. You must use a vectorized function.
4. Create a vector of your three favorite foods. Then, in a separate command, add your favorite drink to
end of the vector.
```{r vector}
# 3way for vector
vec1<- c(-10,-9,-8,-7,-6,-5,-4,-3,-2,-1,0,1,2,3,4,5,6,7,8,9,10)
vec2<- c(-10:10)
vec3<- seq(-10, 21, 1)

# create odd vector 
odd_vec_1 <- c(3,6,9,12,15)
odd_vec_2<- seq(3,17,3)


log10(15:25)
# create vector of food and drink
like_best <- c("Rice","Grill Egg","Vietnamese Tradition Soup")
like_best 
like_best_drk<- append(like_best, "Coffe")
like_best_drk
```

PArt 2
Use vectorized functions for the following:
a. Everyone got 5 points of extra credit on the midterm. Change the midterm grade vector to reflect this.
b. How many students scored higher on the midterm than the final? (Note, use the extra credit version of
the midterm for this)
c. Create a new vector that contains the sum of each student’s midterm and the final exam grade divided
by 2. (You are finding the average for each student)
d. Create a boolean vector that indicates whether each student has an “A” average, greater than or equal
to 90.
e. Use this vector to print/display the names of the students who made an “A”.
f. Challenge (+2 bonus points): How many students made an A if the midterm is 40% of the grade and
the final is 60%.


```{r}
student <- c("a", "b", "c" ,"d" ,"e" ,"f", "g" ,"h", "i" ,"j", "k", "l", "m", "n", "o")
midterm_grades <- c(95, 78, 76, 91, 87, 84, 92, 72, 72, 81, 82, 93, 69, 94, 80)
final_exam_grades <- c(92, 90, 95, 87, 88, 98, 96, 79, 83, 75, 84, 73, 91, 78, 97)

midterm_grades <- midterm_grades+5

midterm_grades

people<- sum(midterm_grades>final_exam_grades)
people

average<- (midterm_grades+final_exam_grades)/2
average


average2<- average>= 90
average2


#index
student_A<- student[average2]

```


Using indexing, and using logical indexing when possible, answer the following questions. Your code
output can be considered an answer.
a. What is the age of the last car?
a. How many green cars are there?
a. How many cars are older than 20 years old?
a. What are the names of the black cars?
a. How many cars had prices greater than 250 or less than 50?
a. Are any of the cars older than 400 years old?
a. Create a new vector that contains the prices of the black and green cars.
a. Check to see if "lambo" is the name of a car.


```{r}
# Data for Part 5
car.names <- c("toyota", "ford", "honda", "bmw", "kia", "dodge", "jeep", "tesla", "subaru", "lexus")
car.colors <- c( "pink", "blue", "blue", "black", "green",
"green", "black", "black", "green", "yellow")
car.ages <- c(13, 5, 36, 2, 4, 67, 24, 2, 6, 16)
car.prices <- c(53, 87, 54, 532, 58, 99, 66, 132, 32, 345)


print_age<- car.ages[length(car.names)]
print_age

corlor<- sum(car.colors=="green") #== assign the value type 
corlor


old_man<- sum(car.ages>20)
old_man


take_name<- car.names[car.colors=="black"]
take_name

up_and_down<- sum(car.prices>250,car.prices<50)
up_and_down


really_old<- car.ages>400
really_old


black_n_green<- c(car.prices[car.colors=="green"],
                  car.prices[car.colors=="black"])
black_n_green

love_luxury<- car.names=="lambo"


car.ages[length(car.names)]



#Part 4
 #list_value<- list(taggednames,colors,ages,prices)

```


Write a function that takes two vectors, finds the maximum value of each vector, and returns the sum of
those maximums. Test the function on two vectors of your own creation.

```{r}
#part 5

two_vector <- function(a,b) {
  max1<-max(a)
  max2<- max(b)
  sum(max1,max2)
}
two_vector(car.ages,car.prices) # making any two vector can work, not assign a fix vector

```

Often you will need to recode values of a dataset. For example, if you have a survey of age data, you may
want to convert any crazy values (like anything below 0 or above 110) to the special value NA.
• Write a function called check.ages() that takes one argument: x, a vector of numbers (and we’ll
assume the user will do so correctly, don’t worry about error checking here). The function should
change any values of x that are below 0 or above 110 to NA, and then return the edited vector.
• Test the function on the vector test_x given below. Your function should return 18 NA 97 NA
test_x <- c(18, -6, 97, 230)

```{r}
#part 6
check.ages<- function(x){

  #filter<- c( x[x > 110] <- "NA",x[x < 0] <- "NA") #not really work
  x[x > 110] <- "NA"
  x[x < 0] <- "NA"
  return(x)
}
test_x <- c(18, -6, 97, 230)

check.ages(test_x)


```


Write a function drone_cost that takes two arguments: the distance of the delivery and whether or
not the customer is a Brime member. The function should return the price of the delivery.
2. Test the function for a Brime member that needs a delivery 14 miles from the warehouse, and 4 different
non-Brime members that are 5, 15,35, and 60 miles from the warehouse .
```{r}
#part 7
#drone_cost Function
drone_cost <- function(distance, brime_member) {
  if (brime_member) {
    cost <- 2.00
  } 
  else if (distance < 0) {
      return("Invalid")
  }
  else if (distance>=0 & distance <= 10 ) {
    cost <- 4.00
  } 
  else if (distance > 10 & distance <= 20) {
    cost <- 7.50
  } 
  else if (distance > 20 & distance <= 50) {
    cost <- 9.00
  } 
  else {
    cost <- 22.50
  }
  return(cost)
}

# Brime member, 14 miles
cat("Cost for Brime member, 14 miles away = $", drone_cost(14, TRUE), "\n")

# Non-Brime members, 5, 15, 35, and 60 miles
cat("Cost for Non-Brime member, 5 miles away = $", drone_cost(5, FALSE), "\n")
cat("Cost for Non-Brime member, 15 miles away = $", drone_cost(15, FALSE), "\n")
cat("Cost for Non-Brime member, 35 miles away = $", drone_cost(35, FALSE), "\n")
cat("Cost for Non-Brime member, 60 miles away = $", drone_cost(60, FALSE), "\n")



```